NGO analysis

### 2022-07-26 Dashboard/NGO Analysis dashboard alpha.Rmd
Copying over report generation andv title box.

@@

### Plots

The dashboard includes five plots, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.

@@

```{r value_boxes, include=FALSE}
# Shiny actions
observe({ # wait for button click

@@

  # creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))
    widen_excel(input$excel$datapath)
  })

@@

@@

  # Creating value box for fundraising ROI
  # Unsure about colour levels
  output$KPI_box_fundraising <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_fundraising_num()),
      caption = "Fundraising ROI",
      # Colour depends on value
      color = if (is.na(KPI_fundraising_num())) {
        "primary"
      } else if (KPI_fundraising_num() < 100) {
        "danger"
      } else if (KPI_fundraising_num() < 200) {
        "warning"
      } else {
        "success"
      },
      icon = if (is.na(KPI_fundraising_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fas fa-donate"
      }
    )
  })

@@

  # Creating value box for OPX KPI
  output$KPI_box_OPX <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_OPX_num()),
      caption = "Liquid assets / obligations",
      color = if (is.na(KPI_OPX_num())) {
        "primary"
      } else if (KPI_OPX_num() < 100) {
        "warning"
      } else {
        "success"
      },
      icon = if (is.na(KPI_OPX_num())) {
        "fa-solid fa-circle-xmark"
      } else if (KPI_OPX_num() < 100) {
        "fa-solid fa-scale-unbalanced"
      } else {
        "fa-solid fa-scale-balanced"
      }
    )
  })

@@

  # Creating value box for private donations growth
  output$KPI_box_private <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_private_num()),
      caption = "Private donation growth",
      color = if (is.na(KPI_private_num())) {
        "primary"
      } else if (KPI_private_num() < 0) {
        "warning"
      } else {
        "success"
      },
      icon = if (is.na(KPI_private_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fas fa-hands-helping"
      }
    )
  })

@@

  # Creating value box for public grants growth
  output$KPI_box_public <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_public_num()),
      caption = "Public grants growth",
      color = if (is.na(KPI_public_num())) {
        "primary"
      } else if (KPI_public_num() < 0) {
        "warning"
      } else {
        "success"
      },
      icon = if (is.na(KPI_public_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fa-solid fa-building-columns"
      }
    )
  })

  ## Public/private proportion--------------------------------------------------
  # Creating public/private proportion
  KPI_proportion <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        transmute(
          proportion =
            sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) /
              sum(revenue_public_grants, revenue_private_contributions,
                revenue_membership_fees,
                na.rm = TRUE
              ) * 100
        ) |>
        pull() |>
        first()
    }
  })

  # Creating value box for private/public proportion
  output$KPI_box_proportion <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_proportion()),
      caption = "Private donations",
      # Adding in the same check as elsewhere - colours match if all are NA
      color = if (is.na(KPI_proportion())) {
        "primary"
      } else {
        "info"
      },
      icon = if (is.na(KPI_proportion())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fa-solid fa-chart-pie"
      }
    )
  })

  ## Excel data test (also below)-----------------------------------------------
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
}) |> bindEvent(input$click)
```

```{r plots, include=FALSE}
# It's necessary to split the observers into separate parts, with all necessary
# code duplicated.

@@

  # creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))
    widen_excel(input$excel$datapath)
  })

>>>>

If desired, a copy of the analysis in PDF format may then be downloaded.

```{r sidebar_button}
# download button
uiOutput("downloadUI")
```

@@

### Graphs

The dashboard includes five graphs, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.

@@

```{r report, include=FALSE}
# Shiny actions begin here =====================================================

# Report Output ================================================================

output$downloadUI <- renderUI({
  downloadButton("report", "Download report")
})

## Download and parameters -----------------------------------------------------
output$report <- downloadHandler(

  filename = "report.pdf",
  content = function(file) {
    tempReport <- file.path(tempdir(), "report.Rmd")
    file.copy("report.Rmd", tempReport, overwrite = TRUE)

    params <- list(
      htmlfile = input$html$datapath,
      excelfile = input$excel$datapath,
      start = input$start,
      end = input$end
    )

    id <- showNotification(
      "Generating report...",
      duration = NULL,
      closeButton = FALSE
    )
    on.exit(removeNotification(id), add = TRUE)

    rmarkdown::render(tempReport,
      output_file = file,
      params = params,
      envir = new.env(parent = globalenv())
    )
  }
)
```

```{r reactive_title_value_boxes, include=FALSE}
# First reactive chunk =========================================================

# wait for button click
observe({

@@

  # creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))
    # Creating data frame
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))
    # Creating data frame
    widen_excel(input$excel$datapath)
  })

  # NGO title ==================================================================
  NGO <- reactive({
    read_excel(input$excel$datapath) |>
      # Turning the needed column into a list
      pull(Firm) |>
      # discarding all NAs in case the name is somewhere
      purrr::discard(is.na) |>
      # choosing the first available name
      first()
  })

  title <- reactive({
    # testing if NGO() is NA
    if (is.na(NGO())) {
      # Printing warning if NA
      "Name of NGO missing"
      # Printing name of NGO
    } else {
      NGO()
    }
  })

  # Creating NGO title using html formatting (colour based on MetBrewer
  # "Hokusai2")
  output$title <- renderText({
    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
  })

@@

  # All boxes use the colour "primary" and the icon fa-solid fa-circle-xmark" if
  # the value is NA

@@

  # Creating value box for fundraising ROI
  # Unsure about colour levels
  output$KPI_box_fundraising <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_fundraising_num()),
      caption = "Fundraising ROI",
      # Colour depends on value
      color = if (is.na(KPI_fundraising_num())) {
        "primary"
      } else if (KPI_fundraising_num() < 100) {
        "danger"
      } else if (KPI_fundraising_num() < 200) {
        "warning"
      } else {
        "success"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_fundraising_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fas fa-donate"
      }
    )
  })

@@

  # Creating value box for OPX KPI
  output$KPI_box_OPX <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_OPX_num()),
      caption = "Liquid assets / obligations",
      # colour depends on value
      color = if (is.na(KPI_OPX_num())) {
        "primary"
      } else if (KPI_OPX_num() < 100) {
        "warning"
      } else {
        "success"
      },
      # icon depends on value
      icon = if (is.na(KPI_OPX_num())) {
        "fa-solid fa-circle-xmark"
      } else if (KPI_OPX_num() < 100) {
        "fa-solid fa-scale-unbalanced"
      } else {
        "fa-solid fa-scale-balanced"
      }
    )
  })

@@

  # Creating value box for private donations growth
  output$KPI_box_private <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_private_num()),
      caption = "Private donation growth",
      # colour depends on value
      color = if (is.na(KPI_private_num())) {
        "primary"
      } else if (KPI_private_num() < 0) {
        "warning"
      } else {
        "success"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_private_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fas fa-hands-helping"
      }
    )
  })

@@

  # Creating value box for public grants growth
  output$KPI_box_public <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_public_num()),
      caption = "Public grants growth",
      # colour depends on value
      color = if (is.na(KPI_public_num())) {
        "primary"
      } else if (KPI_public_num() < 0) {
        "warning"
      } else {
        "success"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_public_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fa-solid fa-building-columns"
      }
    )
  })

  ## Public/private proportion--------------------------------------------------
  # Creating public/private proportion
  KPI_proportion <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # calculating the value as a %
        transmute(
          proportion =
            sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) /
              sum(revenue_public_grants, revenue_private_contributions,
                revenue_membership_fees,
                na.rm = TRUE
              ) * 100
        ) |>
        pull() |>
        # getting the most recent year
        first()
    }
  })

  # Creating value box for private/public proportion
  output$KPI_box_proportion <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_proportion()),
      caption = "Private donations",
      # Adding in the same check as elsewhere - colours match if all are NA
      color = if (is.na(KPI_proportion())) {
        "primary"
      } else {
        "info"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_proportion())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fa-solid fa-chart-pie"
      }
    )
  })

  ## Excel data test -----------------------------------------------------------
  # (also below, in the second chunk)
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
  # the button click
}) |> bindEvent(input$click)
```

```{r reactive_plots, include=FALSE}

# Beginning of the second reactive chunk =======================================
# It's necessary to split the observers into separate parts, with all necessary
# code duplicated, because of R's limitations.

@@

  # Data frames ================================================================
  # Recreating, to make available to this chunk as well

  # creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))
    # Creating data frame
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))
    # Creating data frame
    widen_excel(input$excel$datapath)
  })
